Reducing Acceptance Marks in Emerson-Lei Automata by QBF-Solving

### 2023-03-10 paper/main.tex
few more changes to the paper

TELA have attracted a lot of attention during the last five years.
Their popularity comes probably from the fact that these automata can
often use fewer states than equivalent automata with simpler acceptance
conditions. Further, algorithms handling TELA can automatically handle
all automata with traditional acceptance conditions. TELA can be
obtained for example by translating LTL formulas by \texttt{ltl2dela}
(known as Delag)~\cite{muller.17.gandalf},
\texttt{ltl3tela}~\cite{major.19.atva}, \texttt{ltl2dgra} (known as
Rabinizer~4)~\cite{kretinsky.18.cav}, or \texttt{ltl2tgba} which is a
part of the Spot library~\cite{duret.16.atva2}.  There are also
algorithms processing these automata, for example the emptiness
check~\cite{baier.19.atva} or translation of TELA to parity
automata~\cite{renkin.20.atva,casares.22.tacas}.

@@

\emph{Quantified Boolean formulas (QBF)} are Boolean formulas
extended with universal and existential quantification over
propositional variables. We assume that subformulas of the form
$\forall x.\psi$ and $\exists x.\psi$ do not contain another
quantification of variable $x$ inside $\psi$. The semantics of
$\forall x.\psi$ and $\exists x.\psi$ is given by equivalences
\[
  \begin{array}{rcl}
    \forall x.\psi &~\equiv~& \psi[x\rightarrow\true]~\wedge~\psi[x\rightarrow\false]\\
    \exists x.\psi &~\equiv~& \psi[x\rightarrow\true]~\vee~   \psi[x\rightarrow\false]
  \end{array}
\]
where $\psi[x\rightarrow \rho]$ denotes the formula $\psi$ with all
occurrences of $x$ simultaneously replaced by $\rho$. The equivalences
imply that each QBF can be transformed into an equivalent Boolean
formula. However, the size of this Boolean formula can be exponential
in the size of the original QBF. Let $V$ be the set of all
propositional variables. A mapping $\mu:V\rightarrow\{\true,\false\}$
is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
assignment of an equivalent Boolean formula. A QBF formula is
\emph{satisfiable} iff it has a model.

@@

\begin{table}[b]
\caption{LTL to automata translators used for the experimental evaluation
and the number of successfully translated formulas.}
\label{tab:ltltranslators}
\centering
\begin{tabular}{lr@{~~~~~~}lr}
\toprule
tool        & version&homepage & ~~~automata \\
\midrule
\texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
\texttt{ltl2dgra} & 21.0 & & 234\\
\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
\bottomrule
\end{tabular}
\end{table}

>>>>

TELA have attracted a lot of attention during the last five years.
Their popularity comes probably from the fact that these automata can
often use fewer states than equivalent automata with simpler
acceptance conditions. Further, algorithms handling TELA can
automatically handle all automata with traditional acceptance
conditions. TELA can be obtained for example by translating LTL
formulas with tools \texttt{ltl2dela} (known as
Delag)~\cite{muller.17.gandalf} or \texttt{ltl2dgra} (known as
Rabinizer~4)~\cite{kretinsky.18.cav} of the Owl library,
\texttt{ltl2tgba} of the Spot library~\cite{duret.16.atva2}, or
\texttt{ltl3tela}~\cite{major.19.atva}. There are also algorithms
processing these automata, for example the emptiness
check~\cite{baier.19.atva} or translation of TELA to parity
automata~\cite{renkin.20.atva,casares.22.tacas}.

@@

\emph{Quantified Boolean formulas (QBF)} are Boolean formulas
extended with universal and existential quantification over
propositional variables. We assume that subformulas of the form
$\forall x.\psi$ and $\exists x.\psi$ do not contain another
quantification of variable $x$ inside $\psi$. The semantics of
$\forall x.\psi$ and $\exists x.\psi$ is given by equivalences
\[
  \begin{array}{rcl}
    \forall x.\psi &~\equiv~& \psi[x\rightarrow\true]~\wedge~\psi[x\rightarrow\false]\\
    \exists x.\psi &~\equiv~& \psi[x\rightarrow\true]~\vee~   \psi[x\rightarrow\false]
  \end{array}
\]
where $\psi[x\rightarrow \rho]$ denotes the formula $\psi$ with all
occurrences of $x$ simultaneously replaced by $\rho$. The equivalences
imply that each QBF can be transformed into an equivalent Boolean
formula. However, the size of this Boolean formula can be exponential
in the size of the original QBF. Let $V$ be the set of all
propositional variables. A mapping $\mu:V\rightarrow\{0,1\}$
is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
assignment of an equivalent Boolean formula. A QBF formula is
\emph{satisfiable} iff it has a model.

@@

\begin{table}[b]
\caption{LTL to automata translators used for the experimental evaluation
and the number of successfully translated formulas.}
\label{tab:ltltranslators}
\centering
% \begin{tabular}{lr@{~~~~~~}lr}
% \toprule
% tool        & version&homepage & ~~~automata \\
% \midrule
% \texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
% \texttt{ltl2dgra} & 21.0 & & 234\\
% \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
% \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
% \bottomrule
% \end{tabular}
%
\begin{tabular}{lrlrrr}
  \toprule
  &&&& \multicolumn{2}{c}{automata with} \\ \cmidrule(lr){5-6}
  tool &\!\!\!version&homepage & T/O & 0--1 mark & \textbf{2+ marks}\\
  \midrule
  \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}}
       & ? & ? & \textbf{129} \\
  \texttt{ltl2dgra} & 21.0 &
       & ? & ? & \textbf{234}\\
  \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr}
       & ? & ? & \textbf{70}\\
  \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela}
       & ? & ? & \textbf{91}\\
  \bottomrule
\end{tabular}
\end{table}
